### 2019-01-31 LaTex/Paper_fulltext.tex
Proof-reading

%% 
%% Copyright 2007-2018 Elsevier Ltd
%%

\documentclass[preprint,12pt,authoryear]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
%%\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}
\usepackage{url}
\usepackage{natbib}
%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\journal{Dr. Florian Bader}

\begin{document}

	\begin{frontmatter}

		%% Title, authors and addresses

		%% use the tnoteref command within \title for footnotes;
		%% use the tnotetext command for theassociated footnote;
		%% use the fnref command within \author or \address for footnotes;
		%% use the fntext command for theassociated footnote;
		%% use the corref command within \author for corresponding author footnotes;
		%% use the cortext command for theassociated footnote;
		%% use the ead command for the email address,
		%% and the form \ead[url] for the home page:
		%% use optional labels to link authors explicitly to addresses:
		%% \author[label1,label2]{}
		%% \address[label1]{}
		%% \address[label2]{}

		\title{Visualising Life Satisfaction in Countries across Europe: Individual- and Multi-Level Evidence on Cultural and Pecuniary Factors}

		\author{Nicolas F. K. Göller\fnref{label2}}
		\ead{n.goeller@zeppelin-university.net}
		\fntext[label2]{Student in Sociology, Politics and Economics}

		\author{Noah Peters\fnref{label1}}
		\ead{no.peters@zeppelin-university.net}
		\fntext[label2]{Student in Sociology, Politics and Economics}

		\address{Zeppelin University, Fallenbrunnen 3, 88045 Friedrichshafen\fnref{label3}}
		%% \fntext[label3]{}

		\begin{abstract}
			%% Text of abstract
			This project was set out to provide a comprehensible way for exploring the eclectic determinants of life satisfaction 
			across Europe. An interactive visualisation tool was developed to enable independent inspections of happiness-related
			research questions. While the interactive tool covers the actual analysis and presentation of results, this document
			provides an overview over data and variables used to contribute to contemporary happiness research. 
			The project's aim is not only to spark interest on the determinants of happiness in human life, but also to encourage
			later researchers to critically reflect upon the work presented here in order to point out potential shortcomings. 
			Thus, this document should provide necessary transparency regarding data sourcing and variable selection.
			Lastly, accounts regarding the relatedness of life satisfaction and work-related variables are presented to illustrate one potential
			field of application of our interactive tool. 
		\end{abstract}

	\end{frontmatter}

	%% main text
	\newpage
	\tableofcontents
	\newpage

	\section{Project Scope and Preliminary Remarks}
	Our project investigates life satisfaction across Europe and puts an emphasis on the data visualisation part that comes
	with communicating research results. This aspect appears most relevant with regard to `Open Science' and `Citizen
	Science' approaches that are gaining importance. In order to cater to this motivation, we developed an interactive interface
	- The European Happiness Observer - to inspect life satisfaction and its various antecedents across Europe. The app offers
	a myriad of possibilities to inspect life satisfaction on a descriptive basis or delve deeper by computing regression and multilevel
	analyses. Here, the user can select various options and combinations from a pre-selected pool of variables. \\
	Consciously, the scope of our visualisation tool is kept wider and should steer curiosity and inclination to develop own 
	happiness-related (research) questions. To assure at least some theoretical embeddedness, this paper elaborates upon the
	relation of life satisfaction and work-related variables. This perspective served as a broad orientation during the development
	phase of our tool, illustrates the variable selection and comprehensively exhibits what kind of aspects happiness research
	can elucidate. At this point it should be stressed that the Observer enables the user to examine the questions raised in this
	paper. However, the tool's scope goes far beyond these applications and invites every user to get creative in her own 
	right. Thus, this background paper fulfils two major functions: One the one hand, it communicates the data-driven
	approach of our visualisations i.e., the European Happiness Observer app. On the other hand, it documents the theoretical foundation
	of our variable selection and exemplifies a major question of happiness research, namely the relation to work-related
	variables.

	\section{Motivational Manifesto}
	In a time of widespread digitalisation for many aspects of modern society, 'datafication' and subsequent analysis of data has become ever more prevalent \citep{baack_datafication_2015, lycett_datafication:_2013}. Transparency of data sourcing, data access and methods employed in the course of this analysis is a chief concern of the 'Open Science' movement \citep{delfanti_open_2010}. With more and more transparency, reproducibility of procedures is expected to increase and thus the quality of insights by replication and improvements of methods through a 'wisdom of the (educated) crowd'. There have been extensive developments in the form of guidelines and protocols for researchers, universities and scientific journals that are aimed at promoting more transparency in the scientific process \citep{nosek_promoting_2015, miguel_promoting_2014}. 
	At the same time, large shares of the population lack the essential skills to participate in the research process or perceive the barrier of entry in acquiring the necessary skills to do so, as insurmountably high. The concept of 'Citizen Science' has thus emerged sitting at the edge of both science and education. Citizen science projects enlist non-scientist members of the public to participate in the research (mainly in data collection), allowing to survey large amounts of data, while educating participants about the research topic \citep{shirk_public_2012}. Originally coming from environmental research this approach has already allowed researcher to survey wide ranges of locations and time periods and thus provides access to high-quality data that could quite impossibly be attained by professional research teams on their own \citep{bonney_citizen_2009}.
	Conceptually, 'Open Science' and 'Citizen Science' are deeply intertwined as 'Citizen Science' needs to be as transparent and easily accessible as possible to be effective on a large scale \citep{nov_dusting_2011} and 'Open Science' needs to offer educational value and low entry barriers if it wants to gain attention and participation of a broader audience \citep{newman_future_2012}. We feel committed to the mission of these movements and have thus designed our project in order to advance it for the social sciences.

	\section{Introduction to the Relation of Life Satisfaction and Work-Related Variables}
	Anecdotal evidence suggests that people spend a significant share of their lifetime working (for a casual calculation 
	see: \citet{thompson_what_2016}). Moreover, people's work environment is changing constantly and especially today's 
	discussion about the impact of digitalisation on the labour market requires to investigate how occupational factors shape 
	people's life in general. Since work constitutes a considerable part of people’s life, satisfaction with one’s job and the
	occupational environment must severely determine overall well-being. Drawing on the above-mentioned diversity of 
	occupations, their respective circumstances and adaptation to the ‘digital revolution’, this assumption gains even more 
	ground and relevance: \textit{If work changes, life changes.} \\
	But which work-related factors contribute to a happy - or unhappy - life? Also, how does life satisfaction of the employed
	vary across regions and countries? \\
	Using data form the European Values Study (EVS) and several other sources, we calculated multiple regression models to assess the causality of work-related factors and well-being on the individual level. Moreover, multi-level analyses with regard to social capital and
	macroeconomic as well as developmental indicators were performed to account for variation across regions and countries.

	\section{Theoretical Background and Methodological Implementation}
	Although the causal relation between work and happiness/life satisfaction seems straightforward and intuitively correct,
	its academic investigation is far from trivial. On the one hand, this appears in light of definitional questions regarding the 
	wo variables of interest. On the other hand, operationalising the constructs appropriately remains subject to discussions
	(for subjective well-being see: \citet{kahneman_developments_2006,layard_measuring_2010}). Lastly, the causal nature of job and life
	satisfaction presents itself as far more complex than intuitively assumed. \\
	Each of these three fields is addressed by a comprehensive literature review to set the tone for the subsequent analysis.
	As this analysis draws on specific independent variables on the individual level and several factors on higher aggregate
	levels (multi-level analysis), their respective relation to life satisfaction, as assessed in the literature, is discussed as well.

	\subsection{Life Satisfaction as a Component of Subjective Well-Being (SWB)}
	So far, the terms life satisfaction, happiness, and well-being have been used interchangeably. In fact, however, they
	are commonly understood to act upon each other therefore constituting the wider concept of subjective well-being (SWB)
	that includes life satisfaction, happiness, and positive affect  \citep{bowling_meta-analytic_2010,diener_subjective_1984}. 
	Paying debt to the European Value Study’s (EVS) operationalisation of SWB, we only include life satisfaction as a
	dependent variable: While life satisfaction is measured using a ten-point scale the other available component of SWB,
	happiness, comes in a four-point scale. Consequently, we only included the life satisfaction variable for our regression
	models. For the remains of this study, we concentrate on life satisfaction as a dependent variable.

	\subsection{SWB and Job Satisfaction}
	The literature examining the causal relation of SWB and job satisfaction is extensive
	(for reviews see: \citet{bowling_meta-analytic_2010,erdogan_whistle_2012}). \citet{bowling_meta-analytic_2010}, in their
	meta-analysis, show a positive relationship between all three respective components of SWB and job satisfaction.
	Going even further, \citet{near_job_1987} postulate that 
	\begin{quotation}
	“job satisfaction must logically constitute some part of the
	broader construct of life satisfaction; thus, the question regarding spillover between the two must be ‘how much’ not
	‘whether’." (p. 398). 
	\end{quotation}

	Before we examine the `how much-question', it is necessary to discuss the direction of causality between SWB and job
	satisfaction. In general, the present literature differentiates two approaches: the part-whole theory
	\citep{bowling_meta-analytic_2010,hart_predicting_1999,judge_dispositional_1998} and the dispositional approach. The former
	proposes job satisfaction to act upon SWB. This appears because of the simple reason that has already been mentioned
	above: job satisfaction constitutes a sub-category of overall life satisfaction \citep[cf.]{near_job_1987}.
	Consequently, the approach is also conceived of as a bottom-up approach \citep{erdogan_whistle_2012}. Conversely, the
	dispositional approach assumes a general predisposition of satisfaction with life to affect more specific domains like work
	i.e., satisfaction with those domains \citep{bowling_meta-analytic_2010,diener_subjective_1984,headey_top-down_1991,judge_job_1993,judge_another_1993}. 
	Erdogan et al. (2012) link these predispositions to personal traits that determine how satisfied people feel with certain
	domains and life in general (top-down approach). Positioned between the poles is the spillover hypothesis that just
	contents the basic linkage between job and life satisfaction or even other domains of life 
	\citep{bowling_meta-analytic_2010,erdogan_whistle_2012,judge_job_2001,judge_individual_1994}.
	Conceptually, this relation can be understood as a reciprocal relation between the various domains of life and people’s
	satisfaction (Bowling et al., 2010). \\
	While the direction of causality between life satisfaction and job satisfaction is an open discussion, the interplay of life
	satisfaction with other work-related variables might be easier to grasp. Indeed, our study also examines the effects of
	variables that characterise the working environment beyond perceived job satisfaction. This wider scope appears
	reasonable since 
	\begin{quotation}
	“job satisfaction is only a partial indicator of satisfaction with work domain. A person’s satisfaction with
	career and work-related stress are important elements of the work domain that also need to be understood.” \citep[p. 1042]{erdogan_whistle_2012}.
	\end{quotation}

	\subsection{The Wider Interplay of Work-Related Variables with Life Satisfaction}
Our analysis features a myriad of variables that can be examined with regard to their effect on overall life satisfaction.
In order to justify our variable selection the following remarks elucidate the respective variables' interplay with life satisfaction.
In contrast to the previous examination of the overall concept of SWB, two accounts on the different variables are presented 
in a briefer manner. Rather than covering the entire theoretical background, the following elaboration is though to spur interest
and curiosity to examine the relation of life satisfaction and other factors on one's own - aided by our interactive 
European Happiness Observer interface.

\subsubsection{Life satisfaction and income}
The life satisfaction-income discussion almost always refers back to \citet{easterlin_does_1974} whose observations regarding
the relation of life satisfaction i.e., happiness and income were dubbed the `Easterlin Paradox'. Indeed, his findings from time series data
suggest that life satisfaction remains flat, despite increases in GDP per capita. These findings still heavily inform other researchers'
work and indicate that the relative income in comparison to others might influence SWB \citep{di_tella_happiness_2010,drichoutis_reference_2010,pedersen_happiness_2011}.
Emphasising relative position follows the lead of behavioural labour economics that builds on the findings of prospect theory \citep{kahneman_prospect_1979}.
This concept implies that reference points rather than absolute positions inform people's choices. In the course of behavioural
labour economics, this background is highly relevant \citep{dohmen_behavioral_2014}. 
While investigations examining relative status and changes over time appear highly relevant and theoretically justified, the present
project is not suited to cover these questions. Not being in possession of either time series or relative income data, we need
to focus on absolute income at a certain point in time. Although this independent variable might not capture the underlying relation
of income and life satisfaction, it qualifies as a basic means of inquiry that sheds light on the distinction between life satisfaction
and happiness: “high income buys life satisfaction but not happiness.” \citep[p. 16489]{kahneman_high_2010}.

\subsubsection{Pecuniary vs. non-pecuniary factors}
Income-related investigations also stimulate inquiries examining the opposite of monetary factors, namely non-pecuniary
variables. \textit{Since high income might not be all to strive for, which `softer' factors determine a satisfactory life?}
In general, health takes centre-stage as a determinant of life satisfaction. This statement is not only backed by common 
sense or daily wisdom (Who has not said or been told that health is life's most precious gift?) but also by empirical observations \citep{pedersen_happiness_2011}. \\
In the working environment, health is not a trivial factor either. The \citet{who_global_1994} stresses the importance of health at the work place and terms like `work-life-balance' and `Generation Y' imply changing perceptions of a pleasant working environment. Also \citet{andersson_happiness_2008}
investigates the special relation between self-employment and life satisfaction with ample regard to health issues. \\
Self-employment as such exhibits a variable that is thought to affect life satisfaction \citep{andersson_happiness_2008,schneck_why_2014,van_der_zwan_self-employment_2018}
due to non-pecuniary circumstances that are innate to the status of self-employment. \\
This remark also leads to a greater discussion of features proposed by self-determination theory \citep{deci_intrinsic_1985,deci_self-determination_1989,ryan_self-determination_2000}. \\
Lastly, \citet{luo_essays_2018} uses EVS data to estimate the effects of pecuniary and non-pecuniary factors on life satisfaction
of the unemployed. Intuitively, financial uncertainty might appear as the major concern of unemployment but further investigations
show the significant role that social and psychological factors play. This observation can be interpreted as a general plea for research on
the interplay of non-pecuniary factors of the work environment and life satisfaction.

\subsubsection{Non-pecuniary factors of the work environment}
Drawing on the previous remarks regarding the incorporation of work-related non-pecuniary variables, our analysis comprises of
multiple factors that touch on leadership responsibilities, self-determination related factors, or the occupation's reputation.
The following variables (amongst others) can be examined using the European Happiness Observer:
\begin{itemize}
   \item Work importance in life
   \item Freedom of job decisions
   \item Job development
   \item Work duty to society
   \item Work should come always first
   \item Supervising responsibilities 
   \item Status of employment (i.e., being self-employed)
   \item International Socio-Economic Index of Occupational Status   
\end{itemize}

\subsection{Macro-Level Variables and Multilevel Analyses}
So far, the theoretical background sketched relations between life satisfaction and various individual-level variables. While
this perspective enables us to assess psychological constructs and individual life realities in general, a core feature of our
analysis is the incorporation of multilevel investigations. \\
Being interested in life satisfaction across European countries and regions, we examined variables that might account for 
between-variance. The background on these factors and their respective relation to life satisfaction is examined in this 
section. We point out that the macro-variables cover a wider array than the work-related aspects described above. Nonetheless,
the macro-variables and the multilevel models can be used to supplement individual-level analyses as it is common practice
in multilevel contexts.

\subsubsection{Social capital} 
\citet{bartolini_happy_2014} inspect the relation between life satisfaction and social capital on a micro-level and resume that
social capital can explain variation in life satisfaction in the medium- and long-run while the association is less striking in a
short-run perspective. \citet{mikucka_when_2017} underpin this verdict with their multilevel analysis using World Values Study 
and EVS data. While we did not obtain time-series data to inspect associations over time, we recommend to incorporate 
social capital as a regional level-two variable (here: European NUTS1 regions). Referring to Putnam, \citet{Savelkoul_explaining_2011} imply that social 
capital is especially related to the regional level. \\
Having set the level of inspection, we focussed on the operationalisation of social capital given EVS data. \citet{van_schaik_social_2002}
recommends a procedure specifically related to the EVS that builds on using several categories of social capital as a proxy
for the overall concept. In this fashion, we constructed two indices for (i) institutional trust and (ii) trustworthiness that measures
the prevalence of social norms and legal compliance. Additionally, the EVS includes a binary item covering interpersonal trust. \citet{van_schaik_social_2002}
also associates the respective categories of social capital with European NUTS1 regions, as we do. \\
Comprehensive plots of the proxies can be found in our Observer app and multilevel regressions can be computed.

\subsubsection{Economic macro-data}
Building on \citet{easterlin_does_1974} and the diverse remarks on the association of life satisfaction (see above), GDP presents itself
as a useful multilevel factor. We use country data on GDP for 2008 (the year of the EVS's fourth wave) as a level-3 variable.
Furthermore, Human Development Index (HDI) data can be incorporated to capture a wider socio-economic context.

\subsubsection{Freedom House Ratings}
In order to inspect level-3 effects regarding differences in the political system, we offer to incorporate various Freedom House 
indicators measuring a country's level of democracy. While this perspective offers an additional flavour, it lies not within the main
focus of this project. \\
A detailed description of all data used for this project is presented in the following section.

\section{Data}
All data employed in this project is acquired from official sources and with permission of the provider or in the course of open access. Data as well as original codebooks will be made available for the user upon request, or delivered within the project.
The project is mainly built on data from the 'European Value Study 2008' (EVS), which is the fourth survey wave \citep{evs_european_2016}. The EVS contains questions regarding values of Europeans on aspects like moral, work, family, politics, society, religion, as well as an extensive collection of socio-demographic items. Issues and questions regarding surveying of the data, errata, harmonizing of results etc. may be settled with the publisher or from the original codebook. The file contains a total of 66281 observations for 477 variables over 46 European countries. All transformations on variables can be read in detail from the codebook. Significant changes to the data were made through exclusion of 'Russian Federation' from the sample for the purpose of better visualisation. Furthermore, 'Northern Cyprus' and 'Northern Ireland' which were coded as own countries were added to their respective states after international law. Thus, data for 43 countries was processed for use in the project. Furthermore, aggregations were made from the data in total of 5 cases to create variables that are applicable for multilevel analysis. Finally, a variable of 'Geographical region' was created from the country variable in the EVS. This classification was based on the 'Standard Country or Area Codes for Statistics Use' from the UN Statistics Division \citep{unsd_standard_2019}.
Structural data for nations was obtained from the World Bank for 'GDP per capita of 2008' (in 2018 dollars per inhabitant)\citep{world_bank_gdp_2019}, and for 'Gini Coefficient 2008' (in ration scale 0-1)\citep{world_bank_gini_2019}. In the case of 'Gini Coefficient 2008', missing values for Germany, Poland, Serbia, Kosovo, Macedonia, Croatia and Bosnia Herzegovina were taken from Eurostat 2008 (Germany, Poland) \citep{eurostat_eurostat_2019}, from the CIA World Factbook (Serbia 2008) \citep{central_intelligence_agency_serbia_2019}, from World Bank Gini Coefficient 2009 (Kosovo, Macedonia, Croatia), and from World Bank Gini Coefficient 2007 (Bosnia Herzegovina) \citep{world_bank_gini_2019}. Data was also obtained from the 'Democracy Cross-National' dataset created by Pippa Norris \citep{norris_data_2015} for the Freedom House Democracy rating 2008 and the Human Development Index 2008. Missing values of the Freedom House rating for Kosovo \citep{freedom_house_kosovo_2012} and Montenegro \citep{freedom_house_montenegro_2012} were replaced by the indication for 2008 on the Freedom House website. The missing value of the HDI 2008 for Kosovo was approximated by the value for Serbia in 2008. \\
To obtain national geo data for European countries, the `rnaturalearth' package was used \citep{south_rnaturalearth:_2017}.
Geo data for European regions (NUTS1 2006) were obtained from Eurostat with the according package \citep{lahti_eurostat:_2019}. For the 
processing of geo data, the `sf' package was employed \citep{pebesma_sf:_2018}.

\section{User Manual}
The section is aimed at enabling the user to use the project results competently. To do so, a working model of the statistical software R is required, the following instructions are made for application within its GUI RStudio. The project is divided in 'Data preparation', "Varaible Preparation', 'Visualisation' and 'app'. The latter is the most important for the first time user as it will provide access to the reactive visualisation tool 'European Happiness Observer' that is the centrepiece of this project. First time users, may ignore the other files and can execute the whole 'app' file after installing all required packages. The visualisation tool will open itself subsequently, no further steps or data provision needed. The application has guidelines itself that should be helpful for the user in navigating through its function. Generally speaking, it offers opportunities to build mapplots for the nations and most of the regions at NUTS-1 level, other visualisations can be made using bar charts, bivariate line plots and boxplots grouped by geographical entities. Additionally, it is possible to adapt some linear regression models and display their output. Finally, raw data tables are displayed as the most unrefined view on the data. While the mapplots, bar charts and line plots can be understood quite intuitively, this paragraph is aimed at giving some guidelines for working with the 'Boxplot' and 'Regression table' elements in the application.\\

A boxplot is a visualisation that displays the distribution of data on the y-axis,  The dark bar inside the box indicates the position of the median, while the box itself indicates the value for the first and third quartile of the data. Thus the inside of the box contains 50\% of the data. The antennas around the box (also called 'whiskers') have a distance from it of 1.5 times the size of the box (distance between the third and first quartile, called interquartile range). Any data points outside the range of box and whiskers are considered outliers from the data distribution. Thus, the boxplot allows to compare the spread of data for different groups that are displayed on the x-axis. In the application it is possible, to order the groups after certain variables to see, whether the variable may have an effect on the spread of data.\\

\input{table1}

The regression tables provide the user with the opportunity to compile customized regression models that will give formally statistical evidence on the substance of relations in the data. The simple linear regression without multilevel elements is to be interpreted very similar to the line plot, as starting with the constant value, every increase in the independent variable changes the dependent variable by a certain amount. The number of stars next to the independent variables indicate the significance of its effect (significance as in 'being present'). The total validity of the model can be taken from the value of R$^2$ as it indicates the share of data spread in the dependent variable that is explained by the data spread in the independent variable. In table 1, this is exemplified for the model from 'Job satisfaction' on 'Life satisfaction'.

\input{table2}

However, by introducing multilevel elements into the regression model, the interpretation changes. Multilevel models rest on the assumption that the state of a variable as a characteristic of an individual (like life satisfaction) is expected to be influenced by structural factors. These influences can be accounted for when looking at the effect of variables that are measured on a structural level (like GDP per capita). The effect of this independent variable will then explain differences in the dependent variable that occur because there is variation between the groups, where the structural variable is measured (between-variance). It is then possible to account with variables for the individuals (like job satisfaction) exclusively for the variance inside the group (within-variance). This is exemplified in table 2. Like before, the significance of a variable is indicated by its number of stars. However, the quality of the model calculated is more difficult to assess than before. In table 2 the Akaike Information criterion (AIC) is provided for that purpose. This measure in absolute says nothing about the quality of the model but comparing several model, the one with the smallest value will be the one that explains the most.\\

Users might be interested in seeing more detailed information on variables and can do so in the codebook 'EVS\_final\_cdbk' in the project file. It is also possible to generate the codebook anew by executing the 'Visualisations file'. if users may wish to make changes to data or transformation of variables, the may do so in the files 'Data preparation' and 'Variable preparation'.

\section{Conclusion}
To conclude this project description, it remains to say that the project outline was, at times, torn between several 
different, sometimes conflicting expectations. The ambition to produce high-quality analytical insights combined with 
deep-diving theoretical contextualisation is a challenging endeavour. Conducting a project that allows easily accessible 
data exploration for lay people is tough in its own right. One the one hand, one has to cater to an audience that has only
little understanding of the statistics or programming behing the figures. On the other hand, 
offering enough substance and educational value to spark further interest on the topic and the method should be assured.
These are expectational challenges that are similar to the aforementioned trade-off between 'Open Science' and 'Citizen
Science'. However, the result of this project is promising in so far as it fulfils many of the expectations that it was set out
to meet. The application produced is easily accessible and offers a range of possibilities for the analysis of various
variables on an individuals, regional, and national level. 
Moreover, data sourcing and processing is thoroughly documented. Lastly, significant theoretical considerations
justify our selection of variables and methods and will hopefully attract others to explore the topic of life satisfaction.\\
Future challenges include sufficient data access and quality, especially regarding countries and regions that are not covered to date.
Despite - or maybe because - of future challenges we are optimistic that this project is just the first destination of a longer journey. 
	%% The Appendices part is started with the command \appendix; appendix sections are then done as normal sections
\newpage
 \appendix

\section{Responsibility for Project Areas}

\begin{tabular}{|c|c|c|}
	\hline 
	\textbf{Project area}	& \textbf{Noah Peters} 	& \textbf{Nicolas Goeller} 	\\ 
	\hline 
	Data wrangling			&  		   X			&  				X 			\\ 
	\hline 
	Variable preparation 	&  		   X			&  				X 			\\ 
	\hline 
	Geodata					&  		   X			&  				 			\\ 
	\hline 
	Git	     			 	&  		   				&  				X 			\\ 
	\hline 			
	Indices				 	&  		   X			&  				 			\\ 
	\hline 
	Paper visualisations 	&  		   				&  				X 			\\ 
	\hline 
	Shiny 				 	&  		   				&  				X 			\\ 
	\hline 
	Codebook 			 	&  		   				&  				X 			\\ 
	\hline 
	Geoplots 			 	&  		   X			&  				 			\\ 
	\hline 
	LaTex				 	&  		   X			&  				 			\\ 
	\hline 
	Motivation			 	&  		   				&  				X 			\\ 
	\hline 
	Theory				 	&  		   X			&  				 			\\ 
	\hline 
	User manual			 	&  		   				&  				X 			\\ 
	\hline 
\end{tabular}

	\newpage
	%% If you have bibdatabase file and want bibtex to generate the bibitems, please use
	\bibliographystyle{elsarticle-harv} 
	\bibliography{Literature}

	\newpage
	\hangindent=\parindent
	\parindent 0pt
	\hangafter=-1
	\textbf{Declaration of Authorship}
	\\

	\hangindent=\parindent
	\parindent 0pt
	\hangafter=-1
	We hereby declare that the project paper and all related project elements submitted by the title 
	\\

	\hangindent=\parindent
	\parindent 0pt
	\hangafter=-1
	\textbf{Visualising Life Satisfaction in Countries across Europe: Individual- and Multi-Level Evidence on Cultural and Pecuniary Factors} 
	\\

	\hangindent=\parindent
	\parindent 0pt
	\hangafter=-1
	Are written unaided by ourselves as indicated in the section 'Responsibility for Project Areas'. 
	\\

	\hangindent=\parindent
	\parindent 0pt
	\hangafter=-1
	All indirect or direct sources are acknowledged in the bibliography. We are aware that a false declaration is punishable by law. 
	\\
	\\
	\\
	Place, date                                         	Signature Noah Peters
	\\
	Copenhagen, January 31th 2019\\
	\\
	\\
	\\
	Place, date												Signature Nicolas Goeller
	\\
	Lehrensteinsfeld, January 31th 2019

\end{document}

\endinput
%%